Exploration

### 2018-02-26 exploration.Rmd
fixed establishment probability bug in IPMpack estimates

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
```{r sp_i}
set.seed(20)
subsamp <- 100
mean_abund <- 2000
regPlots <- F
ipmPlots <- F
if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
                    s_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(-1, -.1, 1, -.1), #b1*x1 + ...
                    g_sig=1, # N(E, sig)
                    fl_z=c(-7.5, 1.2, -.2, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-1, -.01, 1, -.01), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z
                    seed_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02))
```

@@

  ## seed production
  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
                               c(X_seed[i,] %*% p$seed_x)))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, 0)
  seed.range <- range(d[[i]]$seed, na.rm=TRUE)

  ## recruits
  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 1)
  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
  d[[i]][rcrt.i,] <- NA
  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])

@@

  ## IPMpack: estimate slopes from d[[i]]
  so <- makeSurvObj(d[[i]], surv ~ size)
  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
  fo <- makeFecObj(d[[i]],
                   Formula=c(fl~size+size2+size3, seed~size),
                   Family=c("binomial", "poisson"),
                   Transform=c("none", "none"))
  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
                         correction="constant")
  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
                          correction="constant")
  IPM2 <- P.mx2 + F.mx2
  lambdas[i] <- Re(eigen(IPM)$values[1])
  lambdas2[i] <- Re(eigen(IPM2)$values[1])

  # Plots
  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.1, col="red")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.1, col="red")
      abline(a=0, b=1, col="blue", lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.1, col="red")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.1, col="red")
    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
    plot(y, stable.dist1, ylab="", type="l", lwd=2,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    lines(y, stable.dist2, lwd=2, lty=2)
    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      xx <- seq(0, hi*.5, length.out=100)
      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
  }
  if(ipmPlots) {
    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
    abline(a=0, b=1, col="white", lty=2)
    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
  }
  if(i %% 100 == 0) cat("Finished cell", i, "\n")
}
lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(2,2))
plot(lambdas ~ temp, data=lam.df, log="y")
plot(lambdas ~ prec, data=lam.df, log="y")
plot(lambdas, lambdas2)
plot(density(log(lambdas))); abline(v=0, col="red", lty=2)
```

>>>>

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
```{r sp_i}
#set.seed(20)
subsamp <- NA
mean_abund <- 2000
regPlots <- F
ipmPlots <- F
if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
                    s_x=c(1, -.1, 1, -.1), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(1, -.1, 1, -.1), #b1*x1 + ...
                    g_sig=1, # N(E, sig)
                    fl_z=c(-3, .6), # b1 + b2*z
                    fl_x=c(1, -.01, 1, -.01), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, 1, -.1), # b1*x1 + ...
                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02))
```

@@

  ## seed production
  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
                               c(X_seed[i,] %*% p$seed_x)))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, NA)
  seed.range <- range(d[[i]]$seed, na.rm=TRUE)

  ## recruits
  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 2)
  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
  d[[i]][rcrt.i,] <- NA
  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])

@@

  ## IPMpack: estimate slopes from d[[i]]
  so <- makeSurvObj(d[[i]], surv ~ size)
  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
  fo <- makeFecObj(d[[i]],
                   Formula=c(fl~size, seed~size+size2),
                   Family=c("binomial", "poisson"),
                   Transform=c("none", "none"),
                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est))
  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
                         correction="constant")
  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
                          correction="constant")
  IPM2 <- P.mx2 + F.mx2

  lambdas[i] <- Re(eigen(IPM)$values[1])
  lambdas2[i] <- Re(eigen(IPM2)$values[1])

  # Plots
  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.2, col="red")
      lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.2, col="red")
      lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.2, col="red")
      lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.2, col="red")
      lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
    yl <- range(c(stable.dist1, stable.dist2))
    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    lines(y, stable.dist2, lwd=2, lty=2, col="blue")
    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      xxx <- seq(0, hi*.5, length.out=100)
      lines(xxx, dnorm(xxx, p$rcrt[1], p$rcrt[2]), col="red")
  }
  if(ipmPlots) {
    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
  }
  if(i %% 100 == 0) cat("Finished cell", i, "\n")
}
lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas2))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(2,2))
plot(log(lambdas) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambdas) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambdas), log(lambdas2)); abline(v=0, h=0, a=0, b=1, lty=c(3,2,2))
plot(density(log(lambdas))); abline(v=0, col="red", lty=2)
```
